---
title: '"Tidy" High-throughout Analysis Data, Examplified by RNA sequencing data'
author: "Tu Hu"
date: "19 April 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
High-throughout analysis revolutionized biology and biochemical sciences. These disciplines were also brought into a big data generation. Such big data was characterized as high-dimensional (consists hundreds and thousands variables), computing and storage resources demanding (difficult for personal computer to handle the analysis tasks).

Therefore, at the first place, when computing and storage resources were expensive and limited. The data storage mechanisms were designed for fitting the performance of "machine", rather than the habits of "human". 


## Old fashioned three-table storage
This essay examplified data "tidying" by a subset of GSE5859 gene expression dataset. GSE5859 is a open-source dataset storing gene expression assay results of a US population study[@GSE5859]. 

This dataset was transformed to R format by Rafael A. Irizarry for tutorial purpose and available on GitHub.

```{r Load Example Data, echo=TRUE, message=FALSE, warning=FALSE}
library(devtools)
install_github("genomicsclass/GSE5859Subset")
library(GSE5859Subset)
data(GSE5859Subset)
```


This dataset consists three tables: "geneAnnotation", "geneExpression", "sampleInfo". 

```{r Dimensions of 3 tables}
dim(geneAnnotation)
dim(geneExpression)
dim(sampleInfo)
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

---
references:
- id: GSE5859
  title: Common genetic variants account for differences in gene expression among ethnic groups
  author:
  - family: Spielman
    given: Richard
  container-title: Nature Genetics
  volume: 39
  URL: 'https://doi.org/10.1038/ng1955'
  DOI: 10.1038/ng1955
  issue: 7
  publisher: Nature Publishing Group
  page: 226-231
  type: article-journal
  issued:
    year: 2007
    month: 1
---

# References
